```{r}
library(ggplot2)
data = read.csv("train.csv")
 
data$Alley = ifelse(is.na(data$Alley), "NoA", data$Alley)
data$BsmtQual = ifelse(is.na(data$BsmtQual), "NB", data$BsmtQual)
data$BsmtCond = ifelse(is.na(data$BsmtCond), "NB", data$BsmtCond)
data$BsmtExposure = ifelse(is.na(data$BsmtExposure), "NB", data$BsmtExposure)
data$BsmtFinType1 = ifelse(is.na(data$BsmtFinType1), "NB", data$BsmtFinType1)
data$BsmtFinType2 = ifelse(is.na(data$BsmtFinType2), "NB", data$BsmtFinType2)
data$FireplaceQu = ifelse(is.na(data$FireplaceQu), "NF", data$FireplaceQu)
data$GarageType = ifelse(is.na(data$GarageType), "NG", data$GarageType)
data$GarageFinish = ifelse(is.na(data$GarageFinish), "NG", data$GarageFinish)
data$GarageQual = ifelse(is.na(data$GarageQual), "NG", data$GarageQual)
data$GarageCond = ifelse(is.na(data$GarageCond), "NG", data$GarageCond)
data$PoolQC = ifelse(is.na(data$PoolQC), "NP", data$PoolQC)
data$Fence = ifelse(is.na(data$Fence), "NF", data$Fence)
data$MiscFeature = ifelse(is.na(data$MiscFeature), "None", data$MiscFeature)

 cols = c('MSSubClass', 'MSZoning', 'Street', 'Alley', 'LotShape', 'LandContour', 
         'Utilities', 'LotConfig', 'LandSlope', 'Neighborhood', 'Condition1', 
         'Condition2', 'BldgType', 'HouseStyle',   
         'RoofStyle', 'RoofMatl', 'Exterior1st', 'Exterior2nd', 'MasVnrType', 
         'ExterQual', 'ExterCond', 'Foundation', 'BsmtQual', 'BsmtCond', 
         'BsmtExposure', 'BsmtFinType1', 'BsmtFinType2', 'Heating', 'HeatingQC', 
         'CentralAir', 'Electrical', 'KitchenQual', 'Functional', 'FireplaceQu', 
         'GarageType', 'GarageFinish', 'GarageQual', 'GarageCond', 'PavedDrive', 
         'PoolQC', 'Fence', 'MiscFeature', 'SaleType', 'SaleCondition', 'YrSold', 'MoSold', 'OverallQual')

 
data$SalePrice = data$SalePrice  / 1000

data[cols] = lapply(data[cols], factor)


# remove  variable
data = subset(data, select = -c(Id))
#  
# data  = subset(data, select = -c(ExterQual,PoolQC, PoolArea, LandSlope,LowQualFinSF, BldgType, MiscFeature, Functional,Utilities))
 

 
 
# remove remaining NA
data = data[complete.cases(data), ]

# save to file
save(data, file="cleaned_data.Rda")


# MODELS
load("cleaned_data.Rda")

```


Data exploration & vizualization  :
```{r}
library(ggridges)
library("tidyverse")
library(hrbrthemes)
library(GGally)
library(viridis)
```

Distribution of sale prices :
```{r}
############### 

ggplot (data, aes ( x=  SalePrice)) + geom_density(fill='#555599') + 
  geom_vline(xintercept = median ( data$SalePrice) , color  = 'red', linewidth= 1.4)+
  geom_text(   y =  0.00004 ,
    x = mean(data$SalePrice) ,
    label = "Median",
    vjust = -1,
    hjust = 1, color='grey' , angle=40
  ) + labs ( x='Sale Price in thousands of dollars' , title= 'Distribution of Sale Prices')+theme_minimal()

```

Sale prices per Neighborhood : 

```{r}
######Sale prices per Neighborhood
data$Neighborhood.mean = 0
grp = data %>%
  group_by(Neighborhood) %>%
  summarise(S  = mean(SalePrice))

 data$Neighborhood.mean = grp$S[ match (data$Neighborhood  , grp$Neighborhood )]

cc <- scales::seq_gradient_pal("blue", "red", "Lab")(seq(0,1,length.out=25))

ggplot ( data, aes ( x=  SalePrice, y= reorder ( Neighborhood,SalePrice ) ,
               group=   Neighborhood,    fill=  Neighborhood.mean  )) + 
  geom_density_ridges(alpha=0.5)  + labs(
  x='Sale Price in thousands of dollars' , y= 'Neighborhood' ,
  title= 'Distribution of  Sale Prices according to Neighborhood'
) +  guides ( fill  = guide_colorbar(title=
                                       'Mean Sale Price'
                                       )) + 
  scale_fill_gradient(high='red', low='green') +  theme_minimal() + theme(plot.title = element_text(size=12))
data = subset(data, select = -c(Neighborhood.mean))

```


Sale prices according to overall quality : 
```{r}

#sale prices according to overallq
ggplot(data, aes(x = as.factor(OverallQual), y = SalePrice, fill= OverallQual)) +
  geom_boxplot(alpha = 0.5) +
  geom_jitter(data= data ,alpha = 0.24  ) +
  labs(
    x =  'Rates the overall material and finish of the house',
    title = 'Sale Prices according to overall quality'
  ) + scale_fill_continuous(high='green', low='red')+
  geom_hline(yintercept = mean(data$SalePrice), color='red', linewidth=1.5) +
  geom_text(
    x = 3,
    y = mean(data$SalePrice),
    label = "Mean Sale Price",
    vjust = -1,
    hjust = 1, color='red'
  ) +  theme_minimal() + theme ( legend.position='none')  

```

```{r}

#sale prices accordng to basement area + height

labels = c("Excellent (100+ inches)	"  , "Good (90-99 inches)"  , 
           "Typical (80-89 inches)"  , "Fair (70-79 inches)" , "Poor (<70 inches")
labels = rev(labels)

ggplot ( data , aes(  x = TotalBsmtSF  ,  y = SalePrice, 
                      ,color= as.integer (BsmtQual) )  ) +
  geom_point() +
  labs (x='Total square feet of basement area', title = 'Sale Prices according to basment area') + 
  scale_color_continuous( low   = '#7AD46C' ,
                          high= 'red', labels=labels)+
  guides ( color=  guide_legend(title= 'Height of the basement', labels= labels) ) +
 theme_minimal() +  theme ( legend.position = c(0.8,0.83))  

```
```{r}

   
  #####sale price according to zoning classification 
  ggplot ( data , aes            (   x=reorder (MSZoning,SalePrice), y=SalePrice, fill= MSZoning )) +
    geom_boxplot() +    scale_x_discrete(
      labels=c(  'C (all)' ='Commercial'   , 'FV'='Floating Village Residential' , 'I'='dfd' 
               , 'RH'='Residential High Density ', 
               'RL' ='Residential Low Density',  
               'RM' =      'Residential Medium Density'      )) +labs (
                 x= 'zoning classification of the sale' , y='Sale Price' , 
                 title = 'Sale Price according to zoning classification  '
               ) +
    scale_colour_hue(l = 45) +  theme_minimal() + 
    theme(legend.position = 'none', axis.text.x = element_text(face="bold", color="#993333",vjust=.97 ,hjust=1,
                                     size=10, angle=45)) 
```
Exploring some one-dominant category variables : 
```{r}

#data2 = subset(data, select = -c(ExterQual,PoolQC, PoolArea, LandSlope,LowQualFinSF, BldgType, MiscFeature, Functional))

 ggplot ( data, aes ( x=  ExterQual)) + geom_bar(fill='lightblue') +
   labs (x= 'quality of the material on the exterior ') + theme_minimal()
```
```{r}

 ggplot ( data, aes ( x=  PoolQC)) + geom_bar(fill='lightblue') +
   labs (x= 'Pool Quality') + theme_minimal()


```
```{r}
table (data$PoolQC)
```
```{r}

 ggplot ( data, aes ( x=  PoolArea)) + geom_bar(fill='lightblue') +
   labs (x= 'PoolArea') + theme_minimal()


```
```{r}

 ggplot ( data, aes ( x=  LandSlope)) + geom_bar(fill='lightblue') +
   labs (x= 'Slope of property') + theme_minimal()


```


Data preparation for Lasso models : 
```{r}
library ( caret)
set.seed(123)
idx <- createDataPartition(data$SalePrice, p = .8, list = FALSE)

train = data[idx, ]
test = data[-idx, ]
 
#Regression preparation:
x_train = model.matrix(SalePrice ~ ., train)
x_test = model.matrix(SalePrice ~ ., test)
y_train = train$SalePrice
y_test = test$SalePrice

```

Lasso regression : 
```{r}
library(glmnet)
set.seed(123)
lasso_model = cv.glmnet(x_train, y_train, alpha = 1)
lambda_lasso = lasso_model$lambda.min
 
lasso_predictions_min = predict(lasso_model, newx = x_test, s = lambda_lasso)
lasso_predictions_1se = predict(lasso_model, newx = x_test, s = lasso_model$lambda.1se )

```


Evaluating the model : 
```{r}

 
MAPE_lasso_min <- mean(abs( (y_test - lasso_predictions_min) / y_test)) * 100
MAPE_lasso_1se <- mean(abs( (y_test - lasso_predictions_1se) / y_test)) * 100

RMSE_lasso_min = sqrt (mean ((y_test - lasso_predictions_min)^2 ))
RMSE_lasso_1se= sqrt (mean ((y_test - lasso_predictions_1se)^2 ))

lasso_coefs_min  = coef ( lasso_model ,   s = "lambda.min" )
lasso_coefs_1se = coef ( lasso_model ,   s = "lambda.1se" )
lasso_variables_min = names (lasso_coefs_min[lasso_coefs_min[ , 1] != 0, ] )
lasso_variables_1se = names (lasso_coefs_1se[lasso_coefs_1se[ , 1] != 0, ] )
 c( MAPE_lasso_min ,MAPE_lasso_1se )

c(RMSE_lasso_min ,RMSE_lasso_1se  )
```


```{r}
#vizualizing the minimal lambda model : 
subtitle = paste0 ( 'MAPE= ' ,round(MAPE_lasso_min,4) , ', RMSE= ',
                round(RMSE_lasso_min,4)     , ' / ', length (lasso_variables_min ), ' variables'  ) 


ggplot ( mapping = aes ( x=y_test  , y=lasso_predictions_min  )) + geom_point( color='red', size=3) +
  geom_abline(intercept = 0, slope=1, color='black')+ labs (
    x= 'Actual House Prices ' , y='Predicted House Prices', 
    title = 'Performance of Lasso model ( Minimal lambda) / Before dropping variables' ,
    subtitle =  subtitle  )+theme(plot.title = element_text(size=10))+
  theme_minimal()
```

```{r}
#vizualizing the 1se lambda model : 
subtitle = paste0 ( 'MAPE= ' ,round(MAPE_lasso_1se,4) , ', RMSE= ',
                round(RMSE_lasso_1se,4)     , ' / ', length (lasso_variables_1se ), ' variables'  ) 


ggplot ( mapping = aes ( x=y_test  , y=lasso_predictions_1se  )) + geom_point( color='red', size=3) +
  geom_abline(intercept = 0, slope=1, color='black')+ labs (
    x= 'Actual House Prices ' , y='Predicted House Prices', 
    title = 'Performance of Lasso model ( 1se lambda) / Before dropping variables' ,
    subtitle =  subtitle  )+theme(plot.title = element_text(size=10))+
  theme_minimal()
 
```
Now let's drop some variables and retrain our model and evaluate it again : 

```{r}
data  = subset(data, select = -c(ExterQual,PoolQC, PoolArea, LandSlope,LowQualFinSF, BldgType, MiscFeature, Functional,Utilities))

```


```{r}
set.seed(123)
idx <- createDataPartition(data$SalePrice, p = .8, list = FALSE)

train = data[idx, ]
test = data[-idx, ]
 
#Regression preparation:
x_train = model.matrix(SalePrice ~ ., train)
x_test = model.matrix(SalePrice ~ ., test)
y_train = train$SalePrice
y_test = test$SalePrice

set.seed(123)
 lasso_model2 = cv.glmnet(x_train, y_train, alpha = 1)
lambda_lasso2 = lasso_model2$lambda.min
 
lasso_predictions2_min = predict(lasso_model2, newx = x_test, s = lambda_lasso2)
lasso_predictions2_1se = predict(lasso_model2, newx = x_test, s = lasso_model2$lambda.1se )

 
MAPE_lasso_min2 <- mean(abs( (y_test - lasso_predictions2_min) / y_test)) * 100
MAPE_lasso_1se2 <- mean(abs( (y_test - lasso_predictions2_1se) / y_test)) * 100

RMSE_lasso_min2 = sqrt (mean ((y_test - lasso_predictions2_min)^2 ))
RMSE_lasso_1se2= sqrt (mean ((y_test - lasso_predictions2_1se)^2 ))

lasso_coefs_min2  = coef ( lasso_model2 ,   s = "lambda.min" )
lasso_coefs_1se2 = coef ( lasso_model2 ,   s = "lambda.1se" )
lasso_variables_min2 = names (lasso_coefs_min2[lasso_coefs_min2[ , 1] != 0, ] )
lasso_variables_1se2 = names (lasso_coefs_1se2[lasso_coefs_1se2[ , 1] != 0, ] )

 c( MAPE_lasso_min2 ,MAPE_lasso_1se2 )

c( RMSE_lasso_min2 , RMSE_lasso_1se2)

```

vizualize again : 
```{r}
#vizualizing the minimal lambda model : 
subtitle = paste0 ( 'MAPE= ' ,round(MAPE_lasso_min2,4) , ', RMSE= ',
                round(RMSE_lasso_min2,4)     , ' / ', length (lasso_variables_min2 ), ' variables'  ) 


ggplot ( mapping = aes ( x=y_test  , y=lasso_predictions2_min  )) + geom_point( color='red') +
  geom_abline(intercept = 0, slope=1, color='black')+ labs (
    x= 'Actual House Prices ' , y='Predicted House Prices', 
    title = 'Performance of Lasso model ( Minimal lambda) / After dropping some variables' ,
    subtitle =  subtitle  )+theme(plot.title = element_text(size=10)) +  theme_minimal()
```

```{r}
#vizualizing the 1se lambda model : 
subtitle = paste0 ( 'MAPE= ' ,round(MAPE_lasso_1se2,4) , ', RMSE= ',
                round(RMSE_lasso_1se2,4)     , ' / ', length (lasso_variables_1se2 ), ' variables'  ) 


ggplot ( mapping = aes ( x=y_test  , y=lasso_predictions2_1se  )) + geom_point( color='red') +
  geom_abline(intercept = 0, slope=1, color='black')+ labs (
    x= 'Actual House Prices ' , y='Predicted House Prices', 
    title = 'Performance of Lasso model ( 1se lambda) / After dropping some variables' ,
    subtitle =  subtitle  )+theme(plot.title = element_text(size=10)) +  theme_minimal()
```

Now will scale the data : 

```{r}
library(caret)
preprocessed <- preProcess(data , method = c("center", "scale"))
scaled_df <- predict(preprocessed, data )
data2   = scaled_df
#splitting again  :
set.seed(123)
idx <- createDataPartition(data2$SalePrice, p = .8, list = FALSE)

train2 = data2[idx, ]
test2 = data2[-idx, ]
 
#Regression preparation:
x_train2 = model.matrix(SalePrice ~ ., train2)
x_test2 = model.matrix(SalePrice ~ ., test2)
y_train2 = train2$SalePrice
y_test2 = test2$SalePrice
 lasso_model_scaled = cv.glmnet(x_train2, y_train2, alpha = 1)
 
 lasso_coefs_1se_scaled  = coef ( lasso_model_scaled ,   s = "lambda.1se" )
 
lasso_coefs_1se_scaled = lasso_coefs_1se_scaled[lasso_coefs_1se_scaled[ , 1] != 0, ]
 

```

```{r}
lasso_coefs_1se_scaled = abs ( lasso_coefs_1se_scaled)
lasso_coefs_1se_scaled=lasso_coefs_1se_scaled[-c(1)]
lasso_coefs_1se = data.frame   (lasso_coefs_1se_scaled) 
 lasso_coefs_1se_scaled$vars = rownames (lasso_coefs_1se_scaled)
 names( lasso_coefs_1se_scaled ) = c('coeffs','vars' )
 
ggplot (lasso_coefs_1se_scaled , aes ( x=abs(coeffs), y=reorder (vars,abs(coeffs) )          )) + geom_point(size=3) +
  labs ( x= 'Absolute value of coefficients after scaling', y='Predictors', 
         title='Variables Importance for Lasso model ( 1se lambda)'
          
           )+ theme_minimal() + theme( plot.title = element_text(size=10))
```

Relaxed Lasso  :

```{r}
set.seed(123)
rl_model = cv.glmnet(x_train, y_train, relax = TRUE)
lambda_rl = rl_model$lambda.min
 
rl_predictions = predict(rl_model, newx = x_test, s = lambda_rl) 
rl_predictions2 =predict(rl_model, newx = x_test, s = rl_model$lambda.1se) 
rl_rmse_min = sqrt(mean((y_test - rl_predictions)^2))
rl_rmse_1se = sqrt(mean((y_test - rl_predictions2)^2))

MAPE_rl_min <- mean(abs( (y_test - rl_predictions) / y_test)) * 100
MAPE_rl_1se <- mean(abs( (y_test - rl_predictions2) / y_test)) * 100

c (rl_rmse_min, rl_rmse_1se )
 c (MAPE_rl_min, MAPE_rl_1se )
```
Final comparision between lasso and relaxed :
```{r}
RMSE_lasso_min2= round (RMSE_lasso_min2,4)
RMSE_lasso_1se2 = round (RMSE_lasso_1se2,4)
rl_rmse_min = round (rl_rmse_min,4)
rl_rmse_1se = round (rl_rmse_1se,4)

final = data.frame (matrix ( ncol = 3))
names ( final) = c ('Model', 'min.lambda' , '1se.lambda' )
 
final = rbind (final , c('Lasso',RMSE_lasso_min2, RMSE_lasso_1se2) )
final = rbind (final , c('Relaxed Lasso' ,rl_rmse_min, rl_rmse_1se) )
final

```

```{r}
set.seed(123)
ridge_model = cv.glmnet(x_train, y_train, alpha=0)
 
 
ridge_predictions_min = predict(ridge_model, newx = x_test, s = ridge_model$lambda.min) 
ridge_predictions_1se =predict(rl_model, newx = x_test, s = rl_model$lambda.1se) 
ridge_rmse_min = sqrt(mean((y_test - ridge_predictions_min)^2))
ridge_rmse_1se = sqrt(mean((y_test - ridge_predictions_1se)^2))

MAPE_ridge_min <- mean(abs( (y_test - ridge_predictions_min) / y_test)) * 100
MAPE_ridge_1se <- mean(abs( (y_test - ridge_predictions_1se) / y_test)) * 100

c (ridge_rmse_min, ridge_rmse_1se )
 c (MAPE_ridge_min, MAPE_ridge_1se )
```

